Notes on conformal manifolds

### 2025-06-23 Draft_MLFlatDirections.tex
Add and correct the pseudo code in Appendix

@@

In this work, we demonstrate the viability of the aformentionned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (1,1)$  supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with some symbolic regression technique. There exist a large litterature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,randall2022bingo,burlacu2019parsimony}, to deep machine learning \cite{petersen2019deep,kamienny2022end}, generative modelling \cite{valipour2021symbolicgpt} and diffusion models \cite{bastiani2025diffusion}. However, in the spirit of trying to build a generalisable method, we develop here our own algorithm, which is based on symbolic regression using an Annealing Importance Sampling method (AIS) \cite{neal1998annealedimportancesampling}. In this paper, we will study a 5 parameter truncation, which is itself coming from a 13 parameter consistent truncation of $\mathcal{N} = (1,1)$ 6-dimensional supergravity.

@@

	The results of this method are as follow. The code finds 8 different following polynomials: 
	\begin{subequations} \label{eq:pols}
    \begin{align}
      p_{1} &= -\sqrt{2}\,x_{1} + \sqrt{2}\,x_{1}x_8 + x_{2}x_8x_{10} - \sqrt{2}\,x_{2}x_{4}x_8,\\
      p_{2} &= 2\,x_{2} - 2\,x_{2}x_8 + \sqrt{2}\,x_{1}x_{10} + 2\,x_{1}x_{4} - x_{2}x_8x_{10}^{2},\\
      p_{3} &= 2\,x_{2} - 2\,x_{2}x_8 + \sqrt{2}\,x_{1}x_8x_{10} + 2\,x_{1}x_{4}x_8
      		- 2\,x_{2} x_{4}^{2}x_8,\\
      p_{4} &= \sqrt{2}\,x_{2} - \sqrt{2}\,x_{2}x_8 + \sqrt{2}\,x_{1}\,x_{4} + x_{1}x_8x_{10}
      		- x_{2}x_{4}x_8x_{10},\\
      p_{5} &= -2\,x_{1} - 2\,x_{2}x_{4}- 2\,x_{1}x_{4}^{2} + 2\,x_{1}x_8 + \sqrt{2}\,x_{2}x_{10}
      		+ x_{1}x_8x_{10}^{2},\\
      p_{6} &= -2\,x_{2}x_{4} - 2\,x_{1}x_{4}^{2} + 2\,x_{2}x_{4}x_8 + \sqrt{2}\,x_{2}x_{10} 
      		- \sqrt{2}\,x_{2}x_8x_{10} +  x_{1}x_8x_{10}^{2},\\
      p_{7} &= -\sqrt{2}\,x_{1}^{2}x_{4} + \sqrt{2}\,x_{2}^{2}x_{4} + \sqrt{2}\,x_{1}x_{2}x_{4}^{2} - x_{1}^{2}x_{10}
      		- x_{2}^{2}x_{10}, \\
      p_{8} &= -2 + 4 x_{8} - 2 x_8^2 + 2 x_{4}^2 x_8 - x_8^2 x_{10}^2
    \end{align}
  \end{subequations}
  where $x_{8} = e^{\tilde{x}_{8}}$. They are found with different frequencies with some polynomials occuring more often than others. The results are presented in \eqref{table:results}.

@@

  In algebraic geometry, the set of solutions to the equations $\nabla V = 0$ defines an algebraic variety over a certain base field. In our case, the potential $V$ and its gradient $\nabla V$ are polynomial functions with
\begin{center}
	\begin{minipage}{16cm}
		\begin{tcolorbox}[
			colback=white,
			colframe=myblue,
			title=\textbf{Algorithme ASMC - Diagramme de flux},
			fonttitle=\sffamily\large,
			halign title=flush center,
			boxrule=2pt,
			left=6pt,
			right=6pt,
			top=8pt,
			bottom=8pt
			]
			\begin{tikzpicture}[node distance=1.6cm and 0.8cm, scale=0.85, transform shape]

@@

			\end{tikzpicture}
		\end{tcolorbox}
	\end{minipage}
\end{center}
%  \begin{center}
%  	\begin{minipage}{18cm}
%  		\begin{tcolorbox}[colback=myblue!5,
%  			colframe=myblue,
%  			title=Algorithm flowchart for ASMC algorithm,
%  			halign title=flush center,
%  			sharp corners
%  			]
%  			%	\scalebox{0.5}{
%  				\begin{tikzpicture}[node distance=1.8cm and .5cm]
%  					
%  					% Nodes
%  					\node (start) [startstop] {Start};
%  					
%  					\node (init_particles) [process, below of=start] {Initialize sparse polynomials};
%  					\node (set_weights) [process, below of=init_particles] {Initialize weights and variables};
%  					\node (compute_beta) [process, below of=set_weights] {Compute or adapt temperature $\beta$};
%  					\node (weight_update) [process, below of=compute_beta] {Compute log-targets and update weights};
%  					
%  					\node (ess_check) [decision, below of=weight_update] {Is ESS $<$ Threshold?};
%  					\node (resample) [process, right=4.25cm of ess_check] {Resample particles};
%  					
%  					\node (mutation) [process, below of=ess_check, yshift=-0.5cm] {Apply MH mutation to particles};
%  					\node (acceptance) [process, below of=mutation] {Compute acceptance and update best};
%  					
%  					\node (end_iter) [decision, below of=acceptance] {Is this the last iteration?};
%  					
%  					\node (exploit) [process, below of=end_iter, yshift=-0.5cm] {Optional local search (exploitation)};
%  					\node (output) [process, below of=exploit] {Return best coefficients and logs};
%  					\node (stop) [startstop, below of=output] {Stop};
%  					
%  					% Arrows
%  					\draw [arrow] (start) -- (init_particles);
%  					\draw [arrow] (init_particles) -- (set_weights);
%  					\draw [arrow] (set_weights) -- (compute_beta);
%  					\draw [arrow] (compute_beta) -- (weight_update);
%  					\draw [arrow] (weight_update) -- (ess_check);
%  					
%  					\draw [arrow] (ess_check) -- node[above] {Yes} (resample);
%  					\draw [arrow] (resample) |- (mutation);
%  					
%  					\draw [arrow] (ess_check) -- node[left] {No} (mutation);
%  					\draw [arrow] (mutation) -- (acceptance);
%  					\draw [arrow] (acceptance) -- (end_iter);
%  					
%  					\draw [arrow] (end_iter) -- node[left] {} ++(-4,0) |- (compute_beta);
%  					\node [pos=10, below=430pt, xshift=-100pt] {No};
%  					\draw [arrow] (end_iter) -- node[right] {Yes} (exploit);
%  					
%  					\draw [arrow] (exploit) -- (output);
%  					\draw [arrow] (output) -- (stop);
%  				\end{tikzpicture}
%  			\end{tcolorbox}
%  		\end{minipage}
%  	\end{center}
  	coefficients lying in a specific field extension of $\mathbb{Q}$—typically of the form $\mathbb{Q}(\sqrt{d_1}, \sqrt{d_2}, \dots)$, where $d_i \in \mathbb{Q}$. That is, the coefficients of the polynomials are combinations of rational numbers and square roots of rational numbers.

@@

We now provide further details about the implementation of our code. To recover the polynomials presented in Eq.~\eqref{eq:pols}, we performed 1000 independent runs, each involving 1000 particles. As can be seen in \eqref{eq:gamman}, we see that what matters the most in the the inverse temperature, but rather the inverse temperature multiplied by the loss function. Therefore, instead of implementing an arbitrary schedule for $\beta$, we employed ans adaptive temperature approach, as previously described. The acceptance ratio was controlled by the following schedule:
\begin{equation}
	\texttt{acceptance\_ratio} = 0.8 - 0.5 \times \left(\frac{\texttt{epoch}}{\texttt{n\_epochs}}\right),
\end{equation}
which encourages exploration during the early stages of training and gradually transitions to a more selective regime. We initialized the algorithm with 
\begin{equation}\label{eq:sample10000adaptativetemp}
	\begin{aligned}
		&\beta = 10^{-6}\,, \\
		&\texttt{p\_{modify} = 0.5},\\
		&\texttt{p\_{multiply} = 0.25},\\
		&\texttt{p\_{divide} = 0.25}. 
	\end{aligned}
\end{equation}
This strategy was determined empirically through trial and error; a systematic analysis of its optimization is left for future work. For the representation of polynomials, we adopt a vectorial approach. Specifically, each polynomial is represented as an array of its coefficients $\{c_k\}$, where $k \in \{1,\dots,n_{\text{mon}}\}$ and $n_{\text{mon}}$ denotes the number of authorized monomials, determined by the chosen maximum degree. We also impose a constraint on the maximum number of terms within a single polynomial to promote sparsity. An alternative approach would be to encourage sparsity through the prior distribution or within the loss function formulation. For the simulations we have done, we chose : 
\begin{equation}
	\begin{aligned}
	&\texttt{max\_degree} = 4,\\
	&\texttt{max\_num\_monomials} = 6.
\end{aligned}
\end{equation}
The main results are presented in Fig.\eqref{fig:evolutionoftargetreproduction} and the losses of the run are shown in Fig. \ref{fig:losses_ASMC}. 
%\begin{figure}
%	\centering
%	\includegraphics[scale=0.35]{Figures/Loss_all_ASMC.png}
%	\caption{Losses of all the 1000 runs in shaded blue and the average loss over all run in brown.}
%	\label{fig:losses_ASMC}
%\end{figure}
The frequencies with which each of the polynomials in Eq.~\eqref{eq:pols} was recovered are reported in the table below:
  \begin{equation}
  	\label{table:results}
	\begin{tabular}{l|ccccccccc}
		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $p_{8}$ &  $\varnothing$  \\\hline\hline
		Frequency & 92.6\% & 75.0\% & 1.7\% & 51.9\% & 0.7\% & 0.1\% & 0.2\% & 0.7\% & 0.3\%
	\end{tabular}
\end{equation}

@@

In this work, we have presented a novel machine learning approach to systematically identify and characterize flat directions in supergravity scalar potentials. Our methodology combines gradient descent sampling with annealed importance sampling for symbolic regression, demonstrating its effectiveness on a 5-scalar subsector of 3d (2,0) supergravity derived from 
$AdS_3 \times S^3 \times T^4$ compactifications of IIB supergravity.

@@

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}

@@

In this work, we demonstrate the viability of the aformentionned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (1,1)$  supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with some symbolic regression technique. There exist a large litterature on symbolic regression, using methods from genetic programming \cite{koza1994genetic} such as \cite{virgolin2021improving,randall2022bingo,burlacu2019parsimony}, to deep machine learning \cite{petersen2019deep,kamienny2022end}, generative modelling \cite{valipour2021symbolicgpt}, diffusion models \cite{bastiani2025diffusion}, and equation learner with the nodes being symbolic operations \cite{2018arXiv180607259S}. Another state-of-the art algorithm is the AIFeynman methods \cite{Udrescu:2019mnk}. They use neural network to identify structure in the dataset (such as translational symmetries, multiplicative separability, compositionality, \dots) to recursively define simpler problems on which they can fit the solutions with polynomials.  However, in the spirit of trying to build a generalisable method, we develop here our own algorithm, which is based on symbolic regression using an Annealing Importance Sampling method (AIS) \cite{neal1998annealedimportancesampling}. In this paper, we will study a 5 parameter truncation, which is itself coming from a 13 parameter consistent truncation of $\mathcal{N} = (1,1)$ 6-dimensional supergravity.

@@

	The results of this method are as follow. The code finds the 8 different following polynomials: 
	\begin{subequations} \label{eq:pols}
    \begin{align}
      p_{1} &= -\sqrt{2}\,x_{1} + \sqrt{2}\,x_{1}x_8 + x_{2}x_8x_{10} - \sqrt{2}\,x_{2}x_{4}x_8,\\
      p_{2} &= 2\,x_{2} - 2\,x_{2}x_8 + \sqrt{2}\,x_{1}x_{10} + 2\,x_{1}x_{4} - x_{2}x_8x_{10}^{2},\\
      p_{3} &= 2\,x_{2} - 2\,x_{2}x_8 + \sqrt{2}\,x_{1}x_8x_{10} + 2\,x_{1}x_{4}x_8
      		- 2\,x_{2} x_{4}^{2}x_8,\\
      p_{4} &= \sqrt{2}\,x_{2} - \sqrt{2}\,x_{2}x_8 + \sqrt{2}\,x_{1}\,x_{4} + x_{1}x_8x_{10}
      		- x_{2}x_{4}x_8x_{10},\\
      p_{5} &= -2\,x_{1} - 2\,x_{2}x_{4}- 2\,x_{1}x_{4}^{2} + 2\,x_{1}x_8 + \sqrt{2}\,x_{2}x_{10}
      		+ x_{1}x_8x_{10}^{2},\\
      p_{6} &= -2\,x_{2}x_{4} - 2\,x_{1}x_{4}^{2} + 2\,x_{2}x_{4}x_8 + \sqrt{2}\,x_{2}x_{10} 
      		- \sqrt{2}\,x_{2}x_8x_{10} +  x_{1}x_8x_{10}^{2},\\
      p_{7} &= -\sqrt{2}\,x_{1}^{2}x_{4} + \sqrt{2}\,x_{2}^{2}x_{4} + \sqrt{2}\,x_{1}x_{2}x_{4}^{2} - x_{1}^{2}x_{10}
      		- x_{2}^{2}x_{10}, \\
      p_{8} &= -2 + 4 x_{8} - 2 x_8^2 + 2 x_{4}^2 x_8 - x_8^2 x_{10}^2
    \end{align}
  \end{subequations}
  where $x_{8} = e^{\tilde{x}_{8}}$. They are found with different frequencies with some polynomials occuring more often than others. The results are presented in \eqref{table:results}.

@@

  In algebraic geometry, the set of solutions to the equations $\nabla V = 0$ defines an algebraic variety over a certain base field. In our case, the potential $V$ and its gradient $\nabla V$ are polynomial  functions with coefficients lying in a specific field extension of $\mathbb{Q}$—typically of the form $\mathbb{Q}(\sqrt{d_1}, \sqrt{d_2}, \dots)$, where $d_i \in \mathbb{Q}$. That is, the coefficients of the polynomials are combinations of rational numbers and square roots of rational numbers.
\begin{figure}[h!]
	\centering
	\begin{minipage}{14cm}
		\begin{tcolorbox}[
			colback=white,
			colframe=myblue,
			title=\textbf{ASMC Algorithm},
			fonttitle=\sffamily\large,
			halign title=flush center,
			boxrule=2pt,
			left=6pt,
			right=6pt,
			top=8pt,
			bottom=8pt
			]
			\begin{tikzpicture}[node distance=1.6cm and 0.8cm, scale=0.85, transform shape]

@@

			\end{tikzpicture}
		\end{tcolorbox}
	\end{minipage}
	\caption{The flow chart of the Annealed Importance Sampling - Sequential Monte Carlo algorithm}
	\label{ASMCflowchart}
\end{figure}

@@

We now provide further details about the implementation of our code. To recover the polynomials presented in Eq.~\eqref{eq:pols}, we performed 1000 independent runs, each involving 1000 particles. As can be seen in \eqref{eq:gamman}, we see that what matters the most in the the inverse temperature, but rather the inverse temperature multiplied by the loss function. Therefore, instead of implementing an arbitrary schedule for $\beta$, we employed ans adaptive temperature approach, as previously described. The acceptance ratio was controlled by the following schedule:
\begin{equation}
	\texttt{acceptance\_ratio} = 0.8 - 0.5 \times \left(\frac{\texttt{epoch}}{\texttt{n\_epochs}}\right),
\end{equation}
which encourages exploration during the early stages of training and gradually transitions to a more selective regime. We initialized the algorithm with 
\begin{equation}\label{eq:sample10000adaptativetemp}
	\begin{aligned}
		&\beta = 10^{-6}\,, \\
		&\texttt{p\_{modify} = 0.5},\\
		&\texttt{p\_{multiply} = 0.25},\\
		&\texttt{p\_{divide} = 0.25}. 
	\end{aligned}
\end{equation}
This strategy was determined empirically through trial and error; a systematic analysis of its optimization is left for future work. For the representation of polynomials, we adopt a vectorial approach. Specifically, each polynomial is represented as an array of its coefficients $\{c_k\}$, where $k \in \{1,\dots,n_{\text{mon}}\}$ and $n_{\text{mon}}$ denotes the number of authorized monomials, determined by the chosen maximum degree. We also impose a constraint on the maximum number of terms within a single polynomial to promote sparsity. An alternative approach would be to encourage sparsity through the prior distribution or within the loss function formulation. For the simulations we have done, we chose : 
\begin{equation}
	\begin{aligned}
	&\texttt{max\_degree} = 4,\\
	&\texttt{max\_num\_monomials} = 6.
\end{aligned}
\end{equation}
The main results are summarized in Fig. \ref{fig:evolutionoftargetreproduction} (a) and (b) and the losses of the run are shown in Fig. \ref{fig:evolutionoftargetreproduction} (c). 
%\begin{figure}
%	\centering
%	\includegraphics[scale=0.35]{Figures/Loss_all_ASMC.png}
%	\caption{Losses of all the 1000 runs in shaded blue and the average loss over all run in brown.}
%	\label{fig:losses_ASMC}
%\end{figure}
The frequencies with which each of the polynomials in Eq.~\eqref{eq:pols} was recovered are reported in the table below:
  \begin{equation}
  	\label{table:results}
	\begin{tabular}{l|ccccccccc}
		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $p_{8}$ &  $\varnothing$  \\\hline\hline
		Frequency & 92.6\% & 75.0\% & 1.7\% & 51.9\% & 0.7\% & 0.1\% & 0.2\% & 0.7\% & 0.3\%
	\end{tabular}
\end{equation}

@@

In this work, we have presented a novel machine learning approach to systematically identify and characterize flat directions in supergravity scalar potentials. Our methodology combines gradient descent sampling with annealed importance sampling for symbolic regression, demonstrating its effectiveness on a 5-scalar subsector of 3d (1,1) supergravity derived from 
$AdS_3 \times S^3 \times T^4$ compactifications of IIB supergravity.

@@

\newpage
\section{Details on the algorithm}

\begin{algorithm}
	\caption{Local Dimension Estimation using PCA}
	\label{alg:local-dim}
	\begin{algorithmic}[1]

		\State \textbf{Function} \textsc{LocalDim1Point}($\texttt{x}$, $\texttt{var\_thres}$)
		\State \hspace{0.5cm} Apply PCA to data matrix $\texttt{x}$
		\State \hspace{0.5cm} Compute explained variance ratios: $\texttt{lambda\_1, lambda\_2,} \ldots, \texttt{lambda\_d}$
		\State \hspace{0.5cm} Compute cumulative variance: $\texttt{cumvar\_k} = \sum_{\texttt{i=1}}^\texttt{k} \texttt{lambda\_i}$ for $\texttt{k = 1, }\ldots\texttt{, d}$
		\State \hspace{0.5cm} Find $\texttt{dim = min\{k : cumvar\_k} \geq \texttt{var\_thres\}}$
		\State \hspace{0.5cm} \textbf{return} $\texttt{dim}$
		\State

%		\State \textbf{Function} \textsc{ComputeLocalDim}(\texttt{i, data, neighbors\_idx, var\_thres})
%		\State \hspace{0.5cm} Extract local neighbors : \texttt{local\_neighbors} $\leftarrow$ \texttt{data[neighbors\_idx[i]]}
%		\State \hspace{0.5cm} \texttt{dim} $\leftarrow$ \textsc{LocalDim1Point}(\texttt{local\_neighbors, var\_thres})
%		\State \hspace{0.5cm} \textbf{return} \texttt{[dim, i]}
%		\State

%		\Require \texttt{data} (input data matrix of size \texttt{n $\times$ d}), \texttt{n\_neig} (number of neighbors), \texttt{var\_thres} (variance threshold)
%		\Ensure Local dimensions for all points

		\State \textbf{Function} \textsc{LocalDimNPoints}(\texttt{data, n\_neig, var\_thres})
		\State \hspace{0.5cm} \texttt{n\_points} $\leftarrow$ number of rows in \texttt{data}
		\State \hspace{0.5cm} Initialize \texttt{results} $\leftarrow \emptyset$
%		\State \hspace{0.5cm} Fit KD-Tree on \texttt{data} with \texttt{n\_neig} neighbors \Comment{Find k-nearest neighbors using KD-Tree}
		\State \hspace{0.5cm} Compute \texttt{neighbors\_idx} for all points in \texttt{data}
		\For{\texttt{i = 1} to \texttt{n\_points}} \Comment{Compute local dimensions for all points}
		\State \texttt{result\_i} $\leftarrow$ \textsc{ComputeLocalDim}(\texttt{i, data, neighbors\_idx, var\_thres})
		\State Append \texttt{result\_i} to \texttt{results}
		\EndFor

		\Return \texttt{results}
	\end{algorithmic}
\end{algorithm}

In this appendix, we detail the different algorithm that we used. We first here provide a pseudo-code for the AIS-SMC algorithm :

%\begin{algorithm}
%	\caption{Annealing Importance Sampling for Polynomial Discovery}
%	\label{alg:annealing-is}
%	\begin{algorithmic}[1]
%		\Require $n\_iter$ (number of iterations), $n\_particles$ (number of particles), $target\_acc\_rate(\cdot,\cdot)$ (target acceptance rate), $adaptation\_strength$ (temperature adaptation parameter)
%		
%		\State Initialize particles $\{z_0^{(k)}\}_{k=1}^{n\_particles}$ with sparse polynomials
%		\State Set $w_0^{(k)} = 1/n\_particles$ and $\tilde{w}_0^{(k)} = 1$ for all $k$ \Comment{normalized and unnormalized weights}
%
%		\State $z_{prev} \leftarrow z_0$, $best\_loss \leftarrow \infty$, $acceptance\_history \leftarrow \emptyset$, $q_{ratio}^{(k)}=1$ for all $k$ \Comment{q is the proposal ratio}, $\beta_0 \leftarrow 10^{-8}$
%		
%		\For{$i = 1$ to $n\_iter$} \Comment{Temperature adaptation}
%
%		\State $\beta_{prev} \leftarrow \beta_{current}$, $avg\_acc \leftarrow$ mean$(acceptance\_history)$ \Comment{only of the last 5 epochs}
%		\If{$avg\_acc < target\_acc\_rate(i, n\_iter) - 0.05$}
%		\State $\beta_{current} \leftarrow \beta_{current} \times (1 - adaptation\_strength)$
%		\ElsIf{$avg\_acc > target\_acc\_rate(i, n\_iter) + 0.05$}
%		\State $\beta_{current} \leftarrow \beta_{current} \times (1 + adaptation\_strength)$
%		\EndIf
%
%		
%		\For{$k = 1$ to $n\_particles$} \Comment{Reweighting step}
%		\State Compute $\log \gamma_i(z_i^{(k)}) = -\beta_{current} \times loss(z_i^{(k)}) $
%		\State Compute $\log \gamma_{i-1}(z_{i-1}^{(k)}) = -\beta_{prev} \times loss(z_{prev}^{(k)})$
%		\State $\Delta \log w_i^{(k)} \leftarrow \log \gamma_i(z_i^{(k)}) - \log \gamma_{i-1}(z_{i-1}^{(k)}) + \log q_{ratio}^{(k)}$
%		\EndFor
%		
%		\State Update unnormalized weights: $\tilde{w}_i^{(k)} \leftarrow \tilde{w}_{i-1}^{(k)} \times \exp(\Delta \log w_i^{(k)})$
%		\State Normalize weights: $w_i^{(k)} \leftarrow \tilde{w}_i^{(k)} / \sum_{j=1}^{n\_particles} \tilde{w}_i^{(j)}$
%		\State Compute ESS: $ESS = 1 / \sum_{k=1}^{n\_particles} (w_i^{(k)})^2$
%		
%		\If{$ESS < n\_particles / 2$} \Comment{Resampling}
%		\State Resample particles: $\{z_i^{(k)}\}_{k=1}^{n\_particles} \sim Multinomial(n\_particles, \{w_i^{(k)}\}_{k=1}^{n\_particles})$
%		\State Reset weights: $w_i^{(k)} = 1/n\_particles$, $\tilde{w}_i^{(k)} = 1$ for all $k$
%		\State Update $z_{prev}$ accordingly
%		\EndIf
%		
%		\State $z_{before\_mutation} \leftarrow z_i$ (save current particles)
%
%		\For{$k = 1$ to $n\_particles$} \Comment{Mutating with Metropolis-Hasting}
%		\State Propose $z'^{(k)} \sim q(z'^{(k)} | z_i^{(k)})$ using specialized MH proposal
%		\State Compute $\log p_{current}^{(k)} = -\beta_{current} \times loss(z_i^{(k)}) + sparsity\_prior(z_i^{(k)})$
%		\State Compute $\log p_{proposed}^{(k)} = -\beta_{current} \times loss(z'^{(k)}) + sparsity\_prior(z'^{(k)})$
%		\State $\log \alpha^{(k)} \leftarrow \log p_{proposed}^{(k)} - \log p_{current}^{(k)} + \log q_{ratio}^{(k)}$
%		\State Generate $u^{(k)} \sim \mathcal{U}(0,1)$
%		\If{$\log u^{(k)} < \log \alpha^{(k)}$}
%		\State $z_i^{(k)} \leftarrow z'^{(k)}$ (accept proposal)
%		\EndIf
%		\EndFor
%		
%		\State $acceptance\_rate \leftarrow$ fraction of accepted proposals
%		\State Add $acceptance\_rate$ to $acceptance\_history$
%		\State $z_{prev} \leftarrow z_{before\_mutation}$ (update for next iteration)
%		
%		\EndFor
%
%		
%		\Return final particles, final losses, final weights
%	\end{algorithmic}
%\end{algorithm}

\begin{algorithm}
	\caption{Annealing Importance Sampling for Polynomial Discovery}
	\label{alg:annealing-is}
	\begin{algorithmic}[1]
		\Require \texttt{n\_iter} (number of iterations), \texttt{n\_particles} (number of particles), \texttt{target\_acc\_rate}$(\cdot,\cdot)$ (target acceptance rate), \texttt{adaptation\_strength} (temperature adaptation parameter)

		\State Initialize particles $\{\texttt{z\_0}^{(k)}\}_{k=1}^{\texttt{n\_particles}}$ with sparse polynomials
		\State Set $\texttt{w\_0}^{\texttt{(k)}} = 1/\texttt{n\_particles}$ and $\texttt{w\_unorm\_0}^{\texttt{(k)}} = 1$ for all $\texttt k$ \Comment{normalized and unnormalized weights}
		\State $\texttt{z\_prev} \leftarrow \texttt{z\_0}$, $\texttt{best\_loss} \leftarrow \infty$, $\texttt{acceptance\_history} \leftarrow \emptyset$, $\texttt{q\_ratio}^{\texttt{(k)}}=1$ for all $k$, $\texttt{beta\_0} \leftarrow \texttt{init\_beta}$ \Comment{q is the proposal ratio}

		\For{$\texttt{i = 1}$ to \texttt{n\_iter}} \Comment{Temperature adaptation}
		\State $\texttt{beta\_{prev}} \leftarrow \texttt{beta\_current}$
		\State $\texttt{avg\_acc} \leftarrow$ mean$(\texttt{acceptance\_history})$ \Comment{only of the last 5 epochs}
		\If{$\texttt{avg\_acc < target\_acc\_rate(i, n\_iter) - 0.05}$}
		\State $\texttt{beta\_current} \leftarrow \texttt{beta\_current} \times \texttt{(1 - adaptation\_strength)}$
		\ElsIf{$\texttt{avg\_acc > target\_acc\_rate(i, n\_iter) + 0.05}$}
		\State $\texttt{beta\_current} \leftarrow \texttt{beta\_current} \times \texttt{(1 + adaptation\_strength)}$
		\EndIf

		\For{$\texttt{k = 1}$ to \texttt{n\_particles}} \Comment{Reweighting step}
		\State Compute $\texttt{log(gamma\_i(z\_i}^{\texttt{(k)}} \texttt{)) = - beta\_current} \times \texttt{loss}(\texttt{z\_i}^{\texttt{(k)}}) $
		\State Compute $\texttt{log(gamma\_(i-1)}(\texttt{z\_(i-1)}^{\texttt{(k)}}\texttt{)) = - beta\_prev} \times \texttt{loss}(\texttt{z\_prev}^{\texttt{(k)}})$
		\State $\Delta \log \texttt{w\_i}^{\texttt{(k)}} \leftarrow \texttt{log(gamma\_i(\texttt{z\_i}}^{\texttt{(k)}}  \texttt{) - log(gamma\_(i-1)(z\_(i-1)}^{\texttt{(k)}} \texttt{) + log(q\_ratio}^{\texttt{(k)}}\texttt)$
		\EndFor

		\State Update unnormalized weights: $\texttt{w\_unorm\_i}^{\texttt{(k)}} \leftarrow \texttt{w\_unorm\_(i-1)}^{\texttt{(k)}} \times \texttt{exp(}\Delta \texttt{log w\_i}^{\texttt{(k)}}\texttt)$
		\State Normalize weights: $\texttt{w\_i}^{\texttt{(k)}} \leftarrow \texttt{w\_unorm\_i}^{\texttt{(k)}} \texttt/ \sum_{\texttt{j=1}}^{\texttt{n\_particles}} \texttt{w\_unorm\_i}^{\texttt{(j)}}$
		\State Compute ESS: $\texttt{ESS = 1 /} \sum_{\texttt{k=1}}^{\texttt{n\_particles}} \texttt{(w\_i}^{\texttt{(k)}})^\texttt2$

		\If{$\texttt{ESS < n\_particles / 2}$} \Comment{Resampling}
		\State Resample particles: $\{\texttt{z\_i}^{\texttt{(k)}}\}_{\texttt{k=1}}^{\texttt{n\_particles}} \sim \texttt{Multinomial}(\texttt{n\_particles}, \{\texttt{w\_i}^{\texttt{(k)}}\}_{\texttt{k=1}}^{\texttt{n\_particles}})$
		\State Reset weights: $\texttt{w\_i}^{\texttt{(k)}} = \texttt{1/n\_particles}$, $\texttt{w\_unorm\_i}^{\texttt{(k)}} \texttt{= 1}$ for all $\texttt k$
		\State Update $\texttt{z\_prev}$ accordingly
		\EndIf

		\State $\texttt{z\_before\_mutation} \leftarrow \texttt{z\_i}$ (save current particles)
		\For{$\texttt{k = 1}$ to \texttt{n\_particles}} \Comment{Mutating with Metropolis-Hasting}
		\State Propose $\texttt{z}'^{\texttt{(k)}} \sim \texttt{q}(\texttt{z}'^{\texttt{(k)}} | \texttt{z\_i}^{\texttt{(k)}})$ using specialized MH proposal
		\State Compute $ \texttt{log(p\_current}^{\texttt{(k)}} \texttt{) = - beta\_current} \times \texttt{loss}(\texttt{z\_i}^{\texttt{(k)}}) + \texttt{sparsity\_prior}(\texttt{z\_i}^{\texttt{(k)}})$
		\State Compute $\texttt{log(p\_proposed}^{\texttt{(k)}} \texttt{) = - beta\_current} \times \texttt{loss}(\texttt{z}'^{\texttt{(k)}}) + \texttt{sparsity\_prior}(\texttt{z}'^{\texttt{(k)}})$
		\State $\texttt{log(alpha}^{\texttt{(k)}} \texttt) \leftarrow  \texttt{log(p\_proposed}^{\texttt{(k)}}  \texttt{) -  log(p\_current}^{\texttt{(k)}}  \texttt{) + log(q\_ratio}^{\texttt{(k)}}\texttt)$
		\State Generate $\texttt{u}^{\texttt{(k)}} \sim \mathcal{U}\texttt{(0,1)}$
		\If{$\texttt{u}^{\texttt{(k)}} \texttt{ < alpha}^{\texttt{(k)}}$}
		\State $\texttt{z\_i}^{\texttt{(k)}} \leftarrow \texttt{z}'^{\texttt{(k)}}$ (accept proposal)
		\EndIf
		\EndFor

		\State $\texttt{acceptance\_rate} \leftarrow$ fraction of accepted proposals
		\State Add $\texttt{acceptance\_rate}$ to $\texttt{acceptance\_history}$
		\State $\texttt{z\_prev} \leftarrow \texttt{z\_before\_mutation}$ (update for next iteration)

		\EndFor

		\Return \texttt{final particles, final losses, final weights}
	\end{algorithmic}
\end{algorithm}

This pseudo code sketch how does our annealing loops work. Here is the pseudo-code for the initialization of each individual polynomial. We chose to generate more terms with a high number of monomials, because in principle there are more possibilities of polynomials with larger number of monomials. This does not prevent the code from finding polynomials with lower number of monomials as we can see on the results. The polynomial that the code find the most often is the polynomial $p_1$ which only has 4 terms instead of 5 or 6 for all others polynomials. The code can find such polynomials by putting gradually one or several of the coefficients to zero. Note also that we could include the possibility in the MH process too add/remove monomials. However, in the spirit of making "small" steps in the search space, we decided not to for now. 
\begin{algorithm}
	\caption{Initialize Sparse Polynomials}
	\label{alg:initialize-sparse}
	\begin{algorithmic}[1]
		\Require $\texttt{n\_particles}$ (number of particles), $\texttt{max\_degree}$ (maximum polynomial degree), $\texttt{num\_vars}$ (number of variables), $\texttt{max\_num\_monomials}$ (maximum number of monomials per polynomial), $\texttt{num\_terms}$ (total number of possible terms)

		\State $\texttt{particles} \leftarrow \mathbf{0}_{\texttt{n\_particles} \times \texttt{num\_terms}}$
		\State $\texttt{max\_monomials} \leftarrow \texttt{max\_num\_monomials}$

		\State $\texttt{n\_mon\_to\_max\_order} \leftarrow \binom{\texttt{max\_degree + num\_vars}}{\texttt{max\_degree}}$ \Comment{Compute total number of possible monomials up to max degree}

		\State $\texttt{n\_mon\_all\_coeff} \leftarrow \emptyset$
		\State $\texttt{n\_tot\_pol} \leftarrow \texttt{0}$ 
		\For{$\texttt{i = 1}$ to $\texttt{max\_num\_monomials}$} \Comment{Compute probability distribution over number of monomials}
		\State $\texttt{n\_mon\_i\_coeff} \leftarrow \binom{\texttt{n\_mon\_to\_max\_order}}{i}$ \Comment{Number of polynomials with exactly $i$ monomials}
		\State $\texttt{n\_tot\_pol} \leftarrow \texttt{n\_tot\_pol + n\_mon\_i\_coeff}$
		\State Append $\texttt{n\_mon\_i\_coeff}$ to $\texttt{n\_mon\_all\_coeff}$
		\EndFor

		\State $\texttt{proba} \leftarrow \texttt{n\_mon\_all\_coeff / n\_tot\_pol}$ \Comment{Normalize to get probabilities}

		\For{$\texttt{i = 1}$ to $\texttt{n\_particles}$} \Comment{Generate particles}

		\State $\texttt{num\_monomials} \sim \texttt{Categorical(\{1, 2, \ldots, max\_monomials\}, proba)}$ \Comment{Sample number of monomials according to computed distribution}

		\State $\texttt{nonzero\_indices}\leftarrow$ sample $\texttt{num\_monomials}$ indices from $\{\texttt{1, 2, }\ldots, \texttt{num\_terms}\}$ without replacement \Comment{Randomly select which monomials will be non-zero}
